# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

The first thing that we want to do is pull the data from teh course website,
unzip the file and import into R as a data table

```{r}
temp <- tempfile()

download.file (paste0 ('https://d396qusza40orc.cloudfront.net/', 
                       'repdata%2Fdata%2Factivity.zip'),
               dest = temp, method = 'curl')

unzip (temp, list = T)

act <- read.csv (unz (temp, 'activity.csv'))

unlink (temp)
```


## Finding the mean and median

Here we look at a histogram of the number of steps taken per day

```{r}
hist (with (act, tapply (steps, date, sum, na.rm = T)), xlab = "Steps per Day",
      main = "Average Steps per Day")
```

We also want to find the mean and median values for steps taken per day.

```{r}
mean (with (act, tapply (steps, date, sum, na.rm = T)), na.rm = T)
median (with (act, tapply (steps, date, sum, na.rm = T)), na.rm = T)
```


## What is the average daily activity pattern?

First we want to average the number of steps in each interval across all days
and plot the results.

```{r}
intStepAvg <- with (act, tapply (steps, interval, mean, na.rm= T))

plot (as.numeric (names (intStepAvg)), intStepAvg, type = 'l', xlab="Interval")

max.int <- intStepAvg  [intStepAvg == max (intStepAvg)]
```

It looks like the interval with the average greatest number of steps is the
`r names (max.int)`th interval with an average of `r max (intStepAvg)` steps.


## Imputing missing values

In order to get rid of missing value bias we will use a single imputation
method using the mean of that time period to impute the values.

```{r}
na.tot <- sum (is.na (act$steps))

actIm <- act

na.r <- is.na (act$steps)

actIm$steps [na.r] <- apply (actIm [na.r,], 1,
                             function (x) mean (actIm [actIm [,3] ==
                                                           gsub(' ','',x [3]),
                                                       'steps'], na.rm = T))
```

It apperas the total number of rows with NA values is `r na.tot`.

We can create a new histogram...
```{r}
hist(with (actIm, tapply(steps, date, sum, na.rm = T)), xlab = "Steps per Day",
      main = "Average Steps per Day (with imputed values)")
```

and check out the new mean and median

```{r}
mean (with (actIm, tapply (steps, date, sum, na.rm = T)), na.rm = T)
median (with (actIm, tapply (steps, date, sum, na.rm = T)), na.rm = T)
```

Our mean and median have both shifted upwards now that we have imputed. This
probably means the values taht we were missing were intervals that had high 
average activty times.

## Are there differences in activity patterns between weekdays and weekends?

We are now going to examine the data and look for differences between weekdays
and weekends.

```{r}
actIm$weekday <- as.factor (grepl ('S', weekdays (as.Date (actIm$date))))
levels (actIm$weekday) <- c('weekday', 'weekend')
```

Here is a graph of that data showing the differences.
```{r}
par (mfrow = c(2,1))

WDintSA <- with (subset (actIm, weekday == "weekday"), 
                 tapply (steps, interval, mean, na.rm= T))

plot (as.numeric (names (WDintSA)), WDintSA, type = 'l', xlab="Interval",
      ylab = 'Average Activity', main = 'Weekday', ylim = c(0,250))

WEintSA <- with (subset (actIm, weekday == "weekend"), 
                 tapply (steps, interval, mean, na.rm= T))

plot (as.numeric (names (WEintSA)), WEintSA, type = 'l', xlab="Interval",
      ylab = 'Average Activity', main = 'Weekend', ylim = c(0,250))
```